---
title: "Example of RescueNet For Forecasting"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

Load libraries.

Put the connection string into the odbcConnect or on windows you can set up
an odbc connection in the odbc manager and just put the connection name below.

```{r}
library(tidyverse)
library(magrittr)
library(lubridate)
library(RODBC)
library(prophet)
cn <- odbcConnect("YOUR DATABASE HERE")
```

Read in data from DB
```{r}
dailyCallVolume <- sqlQuery(cn, "SELECT tdate, count(*) as Incidents
                                FROM RescueNet.Trips
                                GROUP BY tdate")
```

Graph the daily volumes since 2012 until today:
```{r}
dailyCallVolume %>%
  filter(year(ymd(tdate)) >= 201e & ymd(tdate) < Sys.Date()) %>%
  ggplot(aes(x = ymd(tdate), y = Incidents)) +
  geom_line() + 
  ylab("Daily Incidents") +
  xlab("Day") + 
  ggtitle("Daily Incidents")

```

Create a dataset for prophet then fit a basic model with US holidays

```{r}
prophet_volume <- dailyCallVolume %>%
  filter(year(ymd(tdate)) >= 2012 & ymd(tdate) < Sys.Date()) %>%
  select(ds = tdate, y = Incidents)     #Need this because prophet requires 
                                        #date to be 'ds' and forecast to be 'y'

m <- prophet()
m <- add_country_holidays(m, country_name = 'US')
m <- fit.prophet(m, prophet_volume)

```


Make predictions
```{r}
future <- make_future_dataframe(m, periods = 365)
forecast <- predict(m, future)
```
```{r}
plot(m, forecast)
```

Create a dataset with data since 2017.
Get just the upper and lower bounds of the model (and expected value), join with actual values.
```{r}
actual_volumes  <- dailyCallVolume %>%
  filter(year(ymd(tdate)) >= 2017 & ymd(tdate) < Sys.Date()) %>%
  select(ds = tdate, y = Incidents)

actual_volumes$ds %<>%ymd()

forecast_small <- forecast %>%
  select(ds, yhat, yhat_lower, yhat_upper)
forecast_small$ds %<>% ymd()

forcast_actual <- left_join(forecast_small, actual_volumes, by = 'ds')
```

Plot actual values compared with expectations from model:
```{r}
forcast_actual %>%
  filter(ds > '2018-10-01' & ds < '2018-11-01') %>%
  ggplot(aes(x = ds)) +
  geom_ribbon(aes(y = yhat, ymin = yhat_lower, ymax = yhat_upper)) + 
  geom_point( aes(y=y), color = "red", size = 3) +
  geom_line( aes(y=y), color = "red", alpha = 0.5) +
  ylab("Incidents Per Day") +
  xlab("") + 
  ggtitle("Forecasted Incidents vs Actual Incidents")
```

```{r}
forcast_actual %>%
  filter(ds > '2018-12-18' & ds < '2019-01-23') %>%
  ggplot(aes(x = ds)) +
  geom_ribbon(aes(y = yhat, ymin = yhat_lower, ymax = yhat_upper)) + 
  geom_point( aes(y=y), color = "red", size = 3) +
  geom_line( aes(y=y), color = "red", alpha = 0.5) +
  ylab("Incidents Per Day") +
  xlab("Day") + 
  ggtitle("Forecasted Incidents vs Actual Incidents", subtitle =  "Including Holidays")
```

Plot the decomposition graph:
```{r}
prophet_plot_components(m, forecast)
```


Looking to the future:
```{r}
forcast_actual %>%
  filter(ds > '2019-04-05' & ds < '2019-05-01' & ds != Sys.Date()-1) %>% #excluding data from yesterday                                                                              becasuse it's not all available                                                                            at the time I'm running this
  ggplot(aes(x = ds)) +
  geom_ribbon(aes(y = yhat, ymin = yhat_lower, ymax = yhat_upper)) + 
  geom_point( aes(y=y), color = "red", size = 3) +
  geom_line( aes(y=y), color = "red", alpha = 0.5) +
  ylab("Incidents Per Day") +
  xlab("") + 
  ggtitle("Predicting the Future")
```



